---
title: 'Task 2: Sea Lion Pups at Alaska Rookeries'
author: "Elmera Azadpour"
date: "12/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
```

```{r, echo = FALSE, error = FALSE, warning=FALSE, message=FALSE}
sl_pups <- read_csv(here("exam_data","steller_pups.csv"))
sl_pups_clean <- sl_pups %>% clean_names()
```

```{r, echo = FALSE, error = FALSE, warning=FALSE, message=FALSE}
pup_means <- sl_pups_clean %>% 
  group_by(year, sitename) %>% 
  mutate(mean_pupcount = mean(pupcount, na.rm= TRUE))  
```

```{r, echo = FALSE, error = FALSE, warning=FALSE, message=FALSE}
pups_tidy <- pup_means %>% 
  filter(sitename %in% c("Sugarloaf", "Forrester Complex", "Chowiet", "Hazy")) %>% 
  select(sitename, year, mean_pupcount) %>% 
  rename(steller_pups = mean_pupcount)
```

```{r,fig.cap = "**Figure 1**. Mean Steller sea lion pup counts from 1961 - 2015 at five sites from Alaska rookeries. Red correlates to Chowiet. Green correlates to Forrester Complex. Blue correlated to Hazy. Purple correlates to Sugarloaf. Data provided from NOAA National Center for Environmental Information (NCEI).", fig.width= 10}
ggplot(pups_tidy, aes(x = year, 
                      y = steller_pups)) +
  geom_line(aes(col = sitename)) +
  geom_point(aes(col = sitename), 
             alpha = 0.3) +
  theme_minimal() +
  labs(x = "Year", 
       y = "Mean Steller sea lion pup count") +
  scale_x_continuous(breaks=seq(1960,2015, 10)) +
  scale_y_continuous(breaks=seq(0,5500, 1000)) +
  scale_color_discrete("Site name")
```
#### Steller pup conclusions:
+ Site Sugarloadand Chowiest saw the greatest Steller sea lion pup count in from around 165 - 1980, followed by a sharp decline in the following years
+ Site Hazy is steadily increasing over time
+ More recently (2010 - ) site Chowiet shows lowest mean Steller sea lion pup populations.

#### Steller sea lion pup data source: 
- Fritz, Lowell; Sweeney, Kathryn; Lynn, Morgan; Gelatt, Thomas; Gilpatrick, James; Towell, Rodney (2015). Counts of Alaska Steller sea lion pups conducted on rookeries in Alaska from 1961-06-22 to 2015-07-18 (NCEI Accession 0128189). NOAA National Centers for Environmental Information. Dataset. https://doi.org/10.7289/v5862ddr. Accessed 2020-12-04.